Reintroduction of resistant frogs allows recovery in the presence of a lethal fungal disease

### 2023-05-10 doc/manuscript/translocation.qmd
Incorporate coauthor comments into Results and Discussion

## Title: Reintroduction of resistant frogs allows recovery in the presence of a lethal fungal disease

## Alternative title: Evolutionary rescue and reintroduction of resistant frogs allows recovery in the presence of a lethal fungal disease

@@

We conducted a principal component analysis (PCA) of the genomic data to describe the association between sampled populations, and then used two complementary approaches to identify regions of the genome under selection.
First, we used a multivariate linear mixed model to evaluate associations between population type (i.e., naive versus recovering) and specific single nucleotide polymorphisms (SNPs).
Second, we used a splined window analysis to identify larger genomic regions showing differences between population types in *F~ST~* and nucleotide diversity ($\pi_{diff}$ = $\pi_{naive}$ - $\pi_{recovering}$).

@@

Results from the SNP and splined window analyses show that recovering populations have signatures of strong selection on multiple regions of the genome.
The SNP analysis identified 11 outlier SNPs from 7 distinct genes across 4 contigs (@fig-selectionresults B, C: SI, SI Dataset 1).
One of the 7 identified genes did not have an associated annotation.
Frequency differences between the naive and recovering populations ranged from 41% to 86%.
Most of these SNPs showed frequency differences in only a subset of the sampled populations (@fig-allelefrequencies A, B), but the outlier SNP in the RIN3 gene showed consistent differences in frequencies across all populations (@fig-allelefrequencies C).
This is suggestive of strong selection at this locus associated with exposure of frogs to Bd.

@@

In this section, we describe the outcome of *R. sierrae* reintroductions conducted in Yosemite National Park as part of a 15-year effort to reestablish populations of this endangered species (`r min(d1$year)`--`r max(d1$year)`).
Each of the `r n_distinct(d1$site_id, d1$date)` reintroductions involved collection of adult frogs from 1 of `r n_distinct(d1$donor)` recovering, Bd-positive "donor" populations (@fig-yosemap SI) and translocating them to 1 of `r n_distinct(d1$site_id)` nearby recipient sites [@seddon2014].
The donor populations included 2 of the 5 recovering populations in the frog evolution study described above (@fig-allelefrequencies A: population 1 and 4).
Following translocation, adult survival and recruitment of new adults were estimated from CMR surveys and counts of tadpoles and juveniles were obtained from visual encounter surveys (VES).

Estimates of frog survival during the 1 year following translocation indicate that survival was highly variable between recipient sites, but relatively constant within recipient sites (for the subset of sites that received multiple translocations; @fig-translocation-survival).
As such, the average survival of frogs in the first translocation to a site was an important predictor of survival of frogs in subsequent translocations to that site (@fig-transsurvival-postdens SI).
This suggests an important effect of site characteristics and/or (site \* frog) interactions on frog survival.
In addition, 1-year survival was higher for frogs translocated later in the study period than earlier (@tbl-survival-earlylate SI), which we suggest resulted primarily from our improved ability over time to choose recipient sites with higher habitat quality for *R. sierrae*.
Finally, Bd loads were fairly consistent before versus after translocation (@fig-bdload-beforeafter SI), and loads were nearly always well below the level indicative of severe chytridiomycosis and associated frog mortality (5.8 ITS copies on a log~10~ scale) [@vredenburg2010; @joseph2018].

@@

Results from frog translocations indicated that adult survival was often relatively high and that most populations showed evidence of successful reproduction and recruitment (described above).
Although suggestive of population establishment, a decade or more of surveys may be necessary to confirm that populations are in fact self-sustaining [@joseph2018].
To extend our inferences of population establishment beyond those possible from the 2-16 years of site-specific CMR data, we developed a population viability model.
Specifically, to test whether the observed yearly adult survival probabilities in translocated populations were sufficient to promote long-term viability, we built a stage-structured matrix model that captured known frog demography and included demographic and environmental stochasticity.
We parameterized the model using CMR data from translocated populations and known life history values in this system (@tbl-param_values SI).
Given observed yearly adult survival probabilities (from analysis of CMR data) and a baseline yearly recruitment probability $\omega$ of 0.3 (i.e., the probability that, following survival, juveniles successfully recruit to the adult class), our stage-structured model predicted that six of twelve translocated populations should experience a long-run growth rate $\lambda$ greater than 1 in the presence of Bd (@fig-viability A; median predicted $\lambda$ ranges from 1.15-1.36 for these six populations).
These six populations all had observed yearly adult survival greater than 0.5.
When we reduced the probability of successful recruitment $\omega$ to 0.1 (a highly conservative estimate), long-run growth rate of five of these six populations was still greater than 1 (@fig-viability A).
In contrast, for the other six translocated populations, where yearly adult survival was substantially less than 0.5, our model predicted that it was unlikely that long-run growth rate in these populations was greater than 1, even when the probability of successful recruitment to the adult life stage was near 0.4 (@fig-viability A).
A sensitivity analysis of long-run growth rate $\lambda$ to yearly adult survival probability of naturally recruited individuals $\sigma_{A_R}$, recruitment probability $\omega$, and female fecundity *F* showed that $\lambda$ was most sensitive to small changes in $\sigma_{A_R}$ (@fig-viability-supp A).
Note that we assumed $\sigma_{A_R}$ was equivalent to the survival probability of translocated adults $\sigma_{A_T}$, except for three populations where we had \> 5 years of CMR data and recovering populations that together allowed us to estimate $\sigma_{A_R}$ and $\sigma_{A_T}$ separately (@fig-compare_surv_probs SI).
<!--# Consider condensing the above paragraph and providing details in SI. Next paragraph topic sentence could also provide a more direct statement of key results, e.g., something like "Even when incorporating demographic and environmental stochasticity, populations with high adult survival are likely to persist over a 50 year time horizon." -->

@@

Results from our genomic analyses strongly suggest that natural selection for adaptive alleles and resulting increased resistance/tolerance is at least partially responsible for the reduced susceptibility of frogs in recovering populations.
We identified substantial differences in allele frequencies and larger genomic regions between adjacent Bd-naive and recovering MYL frog populations, consistent with strong selection following Bd exposure.
Importantly, some regions under selection are associated with cellular and immunological mechanisms known to contribute to disease resistance, including in amphibians [@zamudio2020].
For example, the MHC plays an important role in immunity.
In our study, we identified a region that shows evidence of selection in recovering populations and contains eight genes associated with either the MHC Class I or Class III regions.
These results corroborate numerous previous studies linking MHC genes to amphibian resistance/tolerance against Bd [e.g., @savage2011; @bataille2015].
Similarly, the region with the strongest indication of directional selection contains the gene, GVINP1.
Several previous studies have found this gene to be differentially expressed during Bd infection [e.g., @grogan2018; @ellison2014].
This gene was also strongly linked to disease in salmon, explaining a remarkable 20% of the resistance phenotype [@robledo2020; @robledo2018].
Finally, the RIN3 gene showed a consistent pattern of allele frequency differences across all nine of the frog populations sampled in this study, indicating consistent selection in populations distributed across a wide geographic area.
In the SI, we provide additional details on these and other genes identified in our study as under selection <!--# Retain or delete? -->.

@@

The recent emergence of Bd worldwide has contributed to the decline of more than 500 amphibian species, nearly 100 of which now are extinct in the wild [@scheele2019].
This unprecedented impact on global amphibian biodiversity is compounded by the lack of any effective and broadly applicable strategies to reverse these impacts [@garner2016; @knapp2022].
Importantly, in addition to the natural recovery documented for MYL frogs [@knapp2016], other species are also showing evidence of post-epizootic recovery in the presence of Bd [@scheele2017; @voyles2018] and suggest the possibility of using animals from these recovering populations to reestablish extirpated populations.
As with MYL frogs, the feasibility and long-term success of such efforts will depend on the availability of robust donor populations containing individuals that have acquired via evolution the adaptive alleles necessary to allow frog survival and population growth in the presence of Bd.
Despite the hopeful example of successful reestablishment of MYL frogs via evolutionary rescue and reintroductions, the challenge of recovering hundreds of Bd-impacted amphibians worldwide is a daunting prospect.
Although we now have a proven strategy to reestablish extirpated MYL frog populations, recovery across their large historical range will require substantial resources over many decades.
The results of this study provide a hopeful starting point for that endeavor, and other future efforts worldwide.

@@

To assess the magnitude of changes in Bd load on frogs following translocation, we compared Bd loads measured before versus after translocation.
Before-translocation loads were quantified using skin swabs collected from all to-be-translocated frogs at the donor site on the day before or the day of the translocation.
After-translocation Bd loads were based on all swabs collected from translocated frogs at the recipient site in the year of and the year following translocation.
Individual frogs and their associated Bd loads were included in the dataset only if frogs were captured at the recipient site at least once during the year 0-1 period.
We used an analysis workflow similar to that described above.
Briefly, the response variable was before-translocation load and the predictor variable was after-translocation load.
We compared models with and without group-level effects, and group-level effects included pit_tag_id, pit_tag_id nested within site_id, or pit_tag_id nested within translocation_id.
All models used a negative binomial family.
Priors, chains, iterations, and checks were as described above.
Code to create the dataset and fit models is available from the following GitHub repository: <https://github.com/SNARL1/translocation>.
<!--# Does it make more sense to only analyze the model with pit_tag_id as the group-level effect? That is the model that is most analogous to a paired t-test. -->

@@

\
<!--# Bree comment: At some point it might be good to talk about what figures go in main text and which stay in supplement (or if we might make some multi-panel figures so that we can fit more in the main text). One thing I'm struck by in the figures is that each of the three main sections (genetics, translocations, PVA) are each so information rich but also somewhat disconnected. Not sure if we want to create some kind of synthetic or hybrid conceptual/data figure that pulls the main point of each analysis out and puts it in visual/conceptual relationship with the others. The results are so cool and we don't want readers to gloss over the details or have to work too hard to put it all back together again. -->

@@

![A population viability analysis of translocated MYL frogs that quantifies the long-run growth rate and extinction probability of frog populations: (A) Predicted long-run growth rate \$\\lambda\$ for different values of yearly adult survival probability \$\\sigma\_{A_R}\$ and probability of recruitment \$\\omega\$, given the parameterized, deterministic model. Colored points show the predicted \$\\lambda\$ values for the twelve translocated populations when probability of recruitment into the adult class is \$\\omega = 0.3\$ (indicated by the dashed line). The red line shows where \$\\lambda = 1\$. (B) Predicted 50-year extinction probabilities of the 12 translocated populations given demographic stochasticity, environmental variability in \$\\omega\$, and different mean values of \$\\omega\$. The dashed line shows where \$\\omega = 0.3\$, for comparison to (A). There are 6 lines at extinction probability = 1, 5 of which (70414--70619) are hidden beneath the line for 70505.](figures/pop_viability_figures_for_manuscript1.jpg){#fig-viability} <!-- Consider a diverging color ramp for panel A, e.g., where lambda=1 is white, lambda>1 is blue, lambda<1 is red -->

@@

![Sensitivity analysis and empirical validation of the stage-structured MYL frog model. (A) Elasticity of $\lambda$ with changes in 3 parameters: $\sigma_{A_R}$ (yearly survival probability of naturally recruited adults), $\omega$ (yearly probability of successful recruitment from juveniles to adults), and *F (*number of eggs produced by a female frog in a year that successfully hatch). Elasticity is calculated at the default parameter values for each population and $\omega$ = 0.3. (B) 100 simulated trajectories (black lines) that most closely matched the observed abundance trajectory of adult amphibians at site 70550 (light orange).](figures/pop_viability_figures_for_supp1.jpg){#fig-viability-supp}

@@

@@

![For frogs translocated to each of the 12 recipient sites, Bd loads for the period immediately prior to translocation versus during the 1-year period after translocation. Box plot details are as in @fig-cond-effects.](figures/bdload_beforeafter.png){#fig-bdload-beforeafter}

@@

| **Parameter**                                                             | **Value**            | **Source**                                                         |
|:-----------------------------|:---------------|:--------------------------|
| $\sigma_{L_1}$, Yearly survival probability of year 1 tadpoles            | 0.7                  | Estimated from field data, observations, natural history knowledge |
| $\sigma_{L_2}$, Yearly survival probability of year 2 tadpoles            | 0.7                  | Estimated from field data, observations, natural history knowledge |
| $\sigma_{L_3}$, Yearly survival probability of year 3 tadpoles            | 0.7                  | Estimated from field data, observations, natural history knowledge |
| $\sigma_{J_1}$, Yearly survival probability of year 1 juveniles           | 0.25                 | Estimated from field data, observations, natural history knowledge |
| $\sigma_{J_2}$, Yearly survival probability of year 2 juveniles           | 0.5                  | Estimated from field data, observations, natural history knowledge |
| $\sigma_{A_R}$, Yearly survival probability of naturally recruited adults | Varies by population | Estimated from mark-recapture studies                              |
| $\sigma_{A_T}$, Yearly survival probability of translocated adults        | Varies by population | Estimated from mark-recapture studies                              |
| $p_{L_1}$, Probability of a year 1 tadpoles remaining as a tadpoles       | 1                    | Estimated from field data, observations, natural history knowledge |
| $p_{L_2}$, Probability of a year 2 tadpoles remaining as a tadpoles       | 0.25                 | Estimated from field data, observations, natural history knowledge |
| $p_{J_1}$, Probability of a year 1 juvenile remaining as a juvenile       | 0.25                 | Estimated from field data, observations, natural history knowledge |
| $p_F$, Probability of a adult female reproducing in a year                | 0.5                  | Could be as high at 1, based on field observations                 |
| $F$, Number of surviving eggs produced by an adult female                 | 100                  | From observations of captive frogs                                 |
| $\omega$, Probability of juvenile successfully recruiting to an adult     | Varies yearly        | Explored different values of this parameter                        |

>>>>

## Title: Evolutionary rescue and reintroduction of resistant frogs allows recovery in the presence of a lethal fungal disease

@@

We conducted a principal component analysis (PCA) of the genomic data to describe the association between sampled populations, and then used two complementary approaches to identify regions of the genome under selection.
First, we used a multivariate linear mixed model to evaluate associations between population type (i.e., naive versus recovering) and specific variants, including single nucleotide polymorphisms (SNPs) and insertions/deletions (INDELS).
Second, we used a splined window analysis to identify larger genomic regions showing differences between population types in *F~ST~* and nucleotide diversity ($\pi_{diff}$ = $\pi_{naive}$ - $\pi_{recovering}$).

@@

Results from the SNP and splined window analyses show that recovering populations have signatures of strong selection on multiple regions of the genome.
The SNP analysis identified 11 outlier SNPs from 7 distinct genes across 4 contigs (@fig-selectionresults B, C: SI, SI Dataset 1).
One of the 7 identified genes did not have an associated annotation.
Frequency differences between the naive and recovering populations ranged from 41% to 86%.
Most of these SNPs showed frequency differences in only a subset of the sampled populations (@fig-allelefrequencies A, B), but the outlier SNP in the RIN3 gene showed consistent differences in frequencies across all populations (@fig-allelefrequencies C).
This is suggestive of strong - and parallel - selection at this locus associated with exposure of frogs to Bd across multiple populations.

@@

To determine whether *R. sierrae* from recovering populations could be used to reestablish extirpated populations, we conducted `r n_distinct(d1$site_id, d1$date)` reintroductions in Yosemite National Park (`r min(d1$year)`--`r max(d1$year)`).
Each of the reintroductions involved collection of adult frogs from 1 of `r n_distinct(d1$donor)` recovering, Bd-positive "donor" populations (@fig-yosemap SI) and translocating them to 1 of `r n_distinct(d1$site_id)` nearby recipient sites [@seddon2014].
The donor populations included 2 of the 5 recovering populations in the frog evolution study described above (@fig-allelefrequencies A: population 1 and 4).
Following translocation, adult survival and recruitment of new adults were estimated from CMR surveys and counts of tadpoles and juveniles were obtained from visual encounter surveys (VES).

Estimates of frog survival during the 1 year following translocation indicate that survival was highly variable between recipient sites, but relatively constant within recipient sites (for the subset of sites that received multiple translocations; @fig-translocation-survival).
As such, the average survival of frogs in the first translocation to a site was an important predictor of survival of frogs in subsequent translocations to that site (@fig-transsurvival-postdens SI).
This suggests an important effect of site characteristics and/or (site \* frog) interactions on frog survival.
In addition, 1-year survival was higher for frogs translocated later in the study period than earlier (@tbl-survival-earlylate SI), which we suggest resulted primarily from our improved ability over time to choose recipient sites with higher habitat quality for *R. sierrae*.
Finally, Bd loads were fairly consistent before versus after translocation and loads were nearly always well below the level indicative of severe chytridiomycosis and associated frog mortality (@fig-bdload-beforeafter SI, see SI text for additional details) [@joseph2018; @vredenburg2010].

@@

Results from frog translocations indicated that adult survival was often relatively high and that most populations showed evidence of successful reproduction and recruitment (described above).
Although suggestive of population establishment, a decade or more of surveys may be necessary to confirm that populations are in fact self-sustaining [@joseph2018].
To extend our inferences of population establishment beyond those possible from the 2-16 years of site-specific CMR data, we developed a population viability model.
Specifically, to test whether the observed yearly adult survival probabilities in translocated populations were sufficient to promote long-term viability, we built a stage-structured matrix model that captured known frog demography and included demographic and environmental stochasticity.
We parameterized the model using CMR data from translocated populations and known life history values in this system (@tbl-param_values SI).

Given observed yearly adult survival probabilities (from analysis of CMR data) and a baseline yearly recruitment probability $\omega$ of 0.3 (i.e., the probability that, following survival, juveniles successfully recruit to the adult class), our stage-structured model predicted that six of twelve translocated populations should experience a long-run growth rate $\lambda$ greater than 1 in the presence of Bd (@fig-viability A; median predicted $\lambda$ ranges from 1.15-1.36 for these six populations).
These six populations all had observed yearly adult survival greater than 0.5.
When we reduced the probability of successful recruitment $\omega$ to 0.1 (a highly conservative estimate), long-run growth rate of five of these six populations was still greater than 1 (@fig-viability A).
In contrast, for the other six translocated populations, where yearly adult survival was substantially less than 0.5, our model predicted that it was unlikely that long-run growth rate in these populations was greater than 1, even when the probability of successful recruitment to the adult life stage was near 0.4 (@fig-viability A).
A sensitivity analysis of long-run growth rate $\lambda$ to yearly adult survival probability of naturally recruited individuals $\sigma_{A_R}$, recruitment probability $\omega$, and female fecundity *F* showed that $\lambda$ was most sensitive to small changes in $\sigma_{A_R}$ (@fig-viability-supp A).
Note that we assumed $\sigma_{A_R}$ was equivalent to the survival probability of translocated adults $\sigma_{A_T}$, except for three populations where we had \> 5 years of CMR data and recovering populations that together allowed us to estimate $\sigma_{A_R}$ and $\sigma_{A_T}$ separately (@fig-compare_surv_probs SI).
<!--# Consider condensing the above paragraph and providing details in SI. Next paragraph topic sentence could also provide a more direct statement of key results, e.g., something like "Even when incorporating demographic and environmental stochasticity, populations with high adult survival are likely to persist over a 50 year time horizon." -->

@@

Results from our genomic analyses strongly suggest that natural selection for adaptive alleles and resulting increased resistance/tolerance is at least partially responsible for the reduced susceptibility of frogs in recovering populations.
We identified specific alleles and genomic regions showing signatures of selection between adjacent Bd-naive and recovering MYL frog populations, consistent with strong selection following Bd exposure.
Importantly, some regions under selection are associated with cellular and immunological mechanisms known to contribute to disease resistance, including in amphibians [@zamudio2020].
For example, the MHC plays an important role in immunity.
In our study, we identified a region that shows evidence of selection in recovering populations and contains eight genes associated with either the MHC Class I or Class III regions.
These results corroborate numerous previous studies linking MHC genes to amphibian resistance/tolerance against Bd [e.g., @savage2011; @bataille2015].
Similarly, the region with the strongest indication of directional selection (as measured by $pi_{diff}$) contains the gene GVINP1.
Several previous studies have found this gene to be differentially expressed during Bd infection [e.g., @grogan2018; @ellison2014].
This gene was also strongly linked to disease in salmon, explaining a remarkable 20% of the resistance phenotype [@robledo2020; @robledo2018].
Finally, the RIN3 gene showed a consistent pattern of allele frequency differences across all nine of the frog populations sampled in this study, indicating consistent selection in populations distributed across a wide geographic area.
In the SI, we provide additional details on these and other genes identified in our study as under selection <!--# Reference complement genes. See Allie's comment in the SI for more on this.  -->.

@@

The recent emergence of Bd worldwide has contributed to the decline of hundreds of amphibian species, some of which are now extinct in the wild [@scheele2019].
This unprecedented impact on global amphibian biodiversity is compounded by the lack of any effective and broadly applicable strategies to reverse these impacts [@garner2016; @knapp2022].
Importantly, in addition to the natural recovery documented for MYL frogs [@knapp2016], other species are also showing evidence of post-epizootic recovery in the presence of Bd [@scheele2017; @voyles2018] and suggest the possibility of using animals from these recovering populations to reestablish extirpated populations.
As with MYL frogs, the feasibility and long-term success of such efforts will depend on the availability of robust donor populations containing individuals that have acquired via evolution the adaptive alleles necessary to allow frog survival and population growth in the presence of Bd.
Despite the hopeful example of successful reestablishment of MYL frogs via evolutionary rescue and reintroductions, the challenge of recovering hundreds of Bd-impacted amphibians worldwide is a daunting prospect.
Although we now have a proven strategy to reestablish extirpated MYL frog populations, recovery across their large historical range will require substantial resources over many decades.
The results of this study provide a hopeful starting point for that endeavor, and other future efforts worldwide.

Evolutionary rescue may be increasingly critical for the survival of wildlife populations in a rapidly changing world.
The possibility - and importance - of evolutionary rescue has been discussed for a wide variety of taxa [e.g., @mills2018; @oziolor2019], and in the specific context of disease-impacted systems [@searle2020].
However, obtaining evidence in wild animal populations for all the causal links in this process (i.e., selection for adaptive alleles $\rightarrow$ increase in host fitness in the presence of the pathogen $\rightarrow$ positive population growth and recovery) is difficult and perhaps impossible.
Nonetheless, we provide a compelling example from the wild of how in the face of a novel stressor evolutionary rescue can produce individuals that harbor adaptive alleles and that these individuals can be used to reestablish extirpated imperiled populations.
We expect that such "assisted evolutionary rescue" will be an essential tool in wildlife conservation in an era of global change.
<!--# Work on this last paragraph, especially the last few sentences. -->

@@

We analyzed skin swabs using standard Bd DNA extraction and qPCR methods [@boyle2004, see SI for details].
To assess the magnitude of changes in Bd load on frogs following translocation, we compared Bd loads measured before versus after translocation.
Before-translocation loads were quantified using skin swabs collected from all to-be-translocated frogs at the donor site on the day before or the day of the translocation.
After-translocation Bd loads were based on all swabs collected from translocated frogs at the recipient site in the year of and the year following translocation.
Individual frogs and their associated Bd loads were included in the dataset only if frogs were captured at the recipient site at least once during the year 0-1 period.
We used an analysis workflow similar to that described above.
Briefly, the response variable was before-translocation load and the predictor variable was after-translocation load.
We compared models with and without group-level effects, and group-level effects included pit_tag_id, pit_tag_id nested within site_id, or pit_tag_id nested within translocation_id.
All models used a negative binomial family.
Priors, chains, iterations, and checks were as described above.
Code to create the dataset and fit models is available from the following GitHub repository: <https://github.com/SNARL1/translocation>.
<!--# Does it make more sense to only analyze the model with pit_tag_id as the group-level effect? That is the model that is most analogous to a paired t-test. --> <!--# Add  -->

@@

@@

![A population viability analysis of translocated MYL frogs that quantifies the long-run growth rate and extinction probability of frog populations: (A) Predicted long-run growth rate \$\\lambda\$ for different values of yearly adult survival probability \$\\sigma\_{A_R}\$ and probability of recruitment \$\\omega\$, given the parameterized, deterministic model. Colored points show the predicted \$\\lambda\$ values for the twelve translocated populations when probability of recruitment into the adult class is \$\\omega = 0.3\$ (indicated by the dashed line). The red line shows where \$\\lambda = 1\$. (B) Predicted 50-year extinction probabilities of the 12 translocated populations given demographic stochasticity, environmental variability in \$\\omega\$, and different mean values of \$\\omega\$. The dashed line shows where \$\\omega = 0.3\$, for comparison to (A). There are 6 lines at extinction probability = 1, 5 of which (70414--70619) are hidden beneath the line for 70505.](figures/pop_viability_figures_for_manuscript.jpg){#fig-viability}

@@

![Sensitivity analysis and empirical validation of the stage-structured MYL frog model. (A) Elasticity of $\lambda$ with changes in 3 parameters: $\sigma_{A_R}$ (yearly survival probability of naturally recruited adults), $\omega$ (yearly probability of successful recruitment from juveniles to adults), and *F (*number of eggs produced by a female frog in a year that successfully hatch). Elasticity is calculated at the default parameter values for each population and $\omega$ = 0.3. (B) 100 simulated trajectories (black lines) that most closely matched the observed abundance trajectory of adult amphibians at site 70550 (light orange).](figures/pop_viability_figures_for_supp.jpg){#fig-viability-supp}

@@

### Frog population recovery

Swab extracts were analyzed singly instead of in triplicate [@kriger2006].
For analysis of swabs collected during 2005--2014, we used standards developed from known concentrations of zoospores [@boyle2004], and after 2014, we used standards based on single ITS1 PCR amplicons [@longo2013].
Based on paired comparisons between samples analyzed using both types of standards, Bd in the study area has an average of 60 ITS1 copies per zoospore.
To express all qPCR results as the number of ITS1 copies, starting quantities obtained using the zoospore standard (measured as "zoospore equivalents") were multiplied by 60.
In addition, all qPCR quantities (regardless of standard) were multiplied by 80 to account for the fact that DNA extracts from swabs were diluted 80-fold during extraction and PCR [@vredenburg2010].

@@

![For frogs translocated to each of the 12 recipient sites, Bd loads for the period immediately prior to translocation versus during the 1-year period after translocation. Bd loads are expressed as the number of ITS1 copies per skin swab, as estimated by qPCR of the Bd ITS1 region. Loads indicative of severe disease are \> 5.8 ITS copies (on a log~10~ scale; see SI text for details). Box plot details are as in @fig-cond-effects.](figures/bdload_beforeafter.png){#fig-bdload-beforeafter}

@@

| **Parameter**                                                             | **Value**            | **Source**                                                         |
|:--------------------------|:------------------|:-------------------------|
| $\sigma_{L_1}$, Yearly survival probability of year 1 tadpoles            | 0.7                  | Estimated from field data, observations, natural history knowledge |
| $\sigma_{L_2}$, Yearly survival probability of year 2 tadpoles            | 0.7                  | Estimated from field data, observations, natural history knowledge |
| $\sigma_{L_3}$, Yearly survival probability of year 3 tadpoles            | 0.7                  | Estimated from field data, observations, natural history knowledge |
| $\sigma_{J_1}$, Yearly survival probability of year 1 juveniles           | 0.25                 | Estimated from field data, observations, natural history knowledge |
| $\sigma_{J_2}$, Yearly survival probability of year 2 juveniles           | 0.5                  | Estimated from field data, observations, natural history knowledge |
| $\sigma_{A_R}$, Yearly survival probability of naturally recruited adults | Varies by population | Estimated from mark-recapture studies                              |
| $\sigma_{A_T}$, Yearly survival probability of translocated adults        | Varies by population | Estimated from mark-recapture studies                              |
| $p_{L_1}$, Probability of a year 1 tadpoles remaining as a tadpoles       | 1                    | Estimated from field data, observations, natural history knowledge |
| $p_{L_2}$, Probability of a year 2 tadpoles remaining as a tadpoles       | 0.25                 | Estimated from field data, observations, natural history knowledge |
| $p_{J_1}$, Probability of a year 1 juvenile remaining as a juvenile       | 0.25                 | Estimated from field data, observations, natural history knowledge |
| $p_F$, Probability of a adult female reproducing in a year                | 0.5                  | Could be as high at 1, based on field observations                 |
| $F$, Number of surviving eggs produced by an adult female                 | 100                  | From observations of captive frogs                                 |
| $\omega$, Probability of juvenile successfully recruiting to an adult     | Varies yearly        | Explored different values of this parameter                        |
